How to compute the fusion product of MV cycles in type A

### 2021-05-15 draft.tex
comments edits to groups section

\section{Introduction}
\label{s:intro}
% 
% The Mirkovi\'c--Vybornov isomorphism
In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov provide a geometric version of symmetric and skew Howe $(\GL_m,\GL_N)$ dualities by relating Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $N\times N$ nilpotent orbits on the one hand 
% conjugacy classes of nilpotent matrices 
and to Nakajima quiver varieties on the other hand. 
This paper generalizes the first of these relations (an isomorphism) to families of slices in the Beilinson--Drinfeld Grassmannian. 
Essential to \cite{mirkovic2007geometric}, this space is used to define a fusion (aka convolution) product on sheaves (representations) but can also be used to define a fusion product on certain cycles (bases). 
% 
This product structure makes the vector space spanned by the Mirkovi\'c--Vilonen cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2019mirkovic}). % conjectured by JEA and proved by BKK
% 
To quote from \cite{mirkovic2007geometric}:
\begin{quotation}
    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}
%
The purpose of this paper is to give a conceptually elementary way to compute this product in type $A$. We do not know how fruitful this work might be in the larger landscape of geometric representation theory as it is plagued by two major limitations: it does not readily generalize outside of type $A$, and it is computationally expensive. 
%

@@

Fix $G = \GL_m$ and let $T\subset G$ be the maximal torus of diagonal matrices. 
% are there other possibilities?? :S 
The coweight lattice of $ G $ is $ \ZZ^m = \Hom(\CC^\times,T)$.
% TODO: Thank you - resolved!
% \acom{This identification assumes we are taking $\CC$ points of $G$ right? i.e.\ it's really $P(T(\CC)) = \Hom(\CC^\times,T(\CC))$ that's $\cong \ZZ^m$ and for instance $P(T(\CC[x]/(x^2))$ will give sth else?} 
% \jcom{No, the coweight lattice is the set of homomorphisms of algebraic groups from $ \mathbb G_m$ to $ GL_m$.  It is independent of the ring/field of definition.}
% 
The \textbf{dominant coweights} of $ G $ 
% is
are the set of $ \lambda = (\lambda_1, \dots, \lambda_m) \in \ZZ^m $ such that $ \lambda_1 \ge \cdots \ge \lambda_m$. 
% 
A coweight $ \mu = (\mu_1, \dots, \mu_m) \in \ZZ^m$ is called \textbf{effective}, if $ \mu_j \ge 0 $ for all $ j$.
% We also need the notion of effective dominant coweight, 
% \acom{can we conflate or should we be precise and say coweight?} \jcom{Corrected, let's stick to coweights.}
%which means those $ \lambda $ as above with $ \lambda_m \ge 0$.  
% \acom{1. No boldface on this new term. Proceed with selective boldfacing? 2. Since we sometimes need $\Peff$ which contains $P_{++} = \Peff\cap\Pdom$, how about just using the notation $\Peff,\Pdom$ instead of $P_+, P_{++}$ for a change?}
% \jcom{I added the definition of effective coweight.  I actually don't think that we need any special notation $ P, P_+, P_{dom}$ etc !  Later we can just write ``$ \lambda \in \ZZ^m$ a dominant or effective coweight'' etc}
% TODO: it's settled then, no special notation fro coweights!
% 
So an effective dominant coweight $ \lambda $ is a partition of $|\lambda| = \lambda_1 + \cdots + \lambda_m\in\NN$, which we call the size of $ \lambda$.
%We write $P_+$ for the set of dominant coweights, and $P_{++}$ for the set of effective dominant coweights. 
% 
% nicematrix test
% $\begin{bNiceArray}{ccc|c}[margin]
%     \Block{3-3}<\Large>{A} & & & 0 \\
%     & \hspace*{1cm} & & \Vdots \\
%     & & & 0 \\
%     \hline
%     0 & \Cdots& 0 & 0
%     \end{bNiceArray}$
% 
% $\begin{pNiceMatrix}
%     \frac12 & -\frac12 \\
%     \frac13 & \frac14 \\
%     \end{pNiceMatrix}$
% 
% Set $\AA = \AA^1_x$ (the subscript denoting the chosen local coordinate) and fix $s\in \AA - \{0\}$. 
%

Given a coweight $\mu \in \ZZ^m$ and a point $ s\in \CC$, we define $ (t-s)^\mu$ to be the diagonal matrix 
% with entries $ (t-s)^{\mu_1}, \dots, (t-s)^{\mu_m}$
% \[
%     \begin{pNiceMatrix}[nullify-dots,xdots/line-style=loosely dotted]
%         (t-s)^{\mu_1} & 0             & \Cdots & 0 \\
%         0             & (t-s)^{\mu_2} & \Ddots & \Vdots \\
%         % 0 & b & a & \Ddots & & \\
%         \Vdots              & \Ddots        & \Ddots & 0 \\
%         % \Vdots & & & & \\
%         0             & \Cdots        &   0    & (t-s)^{\mu_m}
%     \end{pNiceMatrix}
% \]
\[
\begin{bmatrix}
    (t-s)^{\mu_1} \\
    & (t-s)^{\mu_2} \\ 
    & & \ddots \\
    & & & (t-s)^{\mu_m}
\end{bmatrix} 
\]
which we will view in 
% $\GL_m(K) $ 
$G(K)$ for any ring $ K $ containing $\CC[t]$ in which $t-s$ is invertible (for example $ K = \Ks, \CC(t)$). % A: Think R sb K; changed it 
\acom{Add warning that we will also sometimes write $(t-s)^\mu$ for its image in 
% $\GL_m(K)/\GL_m(R)$ 
$G(K)/G(R)$ and we hope it will be clear? Only we haven't introduced this space yet.}
% Resolved
% 2. Change notation $\mu = (\mu_1,\dots,\mu_m)$ since we also have $\mu = \mu_1 + \mu_2$. We could do $\mu = \mu' + \mu''$ or $\mu = a + b$? The notation $\mu^{(i)}$ is already taken to mean sth. Ofc all can be changed. Can also do $\mu = {}_{1}\mu + {}_{2}\mu$. \emoji{sweat-smile}}

We will also be interested in the affine space $ M_m$ of $m\times m$ matrices.  Note that for any $ \CC$-algebra $ R $, 
% $\GL_m(R) $ 
$G(R)$ consists of those matrices $ g \in M_m(R) $ whose determinant is invertible in $ R$. Thus for example $ (t-s)^\mu\in M_m(\CC[t])$ for all effective $ \mu \in \ZZ^m$ but 
% $ t^\lambda \in \GL_m(\CC[t]) $ 
$(t-s)^\mu\in G(\CC[t])$ if and only if $ \mu = 0 $.
% 
\acom{maybe add that it's the same for $G(\cO)$.}

>>>>

\section{Introduction}
\label{s:intro}
% 
Throughout this paper (unless indicated otherwise) the base field is $\CC$.

% The Mirkovi\'c--Vybornov isomorphism
In \cite{mirkovic2007quiver} (see also the recent sequel \cite{mirkovic2019comparison}) Mirkovi\'c and Vybornov provide a geometric version of symmetric and skew Howe $(\GL_m,\GL_N)$ dualities by relating Kazhdan--Lusztig slices in the affine Grassmannian of $\GL_m$ to slices in $N\times N$ nilpotent orbits on the one hand 
% conjugacy classes of nilpotent matrices 
and to Nakajima quiver varieties on the other hand. 
This paper generalizes the first of these relations (an isomorphism) to families of slices in the Beilinson--Drinfeld Grassmannian. 
Essential to \cite{mirkovic2007geometric}, this space is used to define a fusion (aka convolution) product on sheaves (representations) but can also be used to define a fusion product on certain cycles (bases). 
% 
This product structure makes the vector space spanned by the Mirkovi\'c--Vilonen cycles into a commutative algebra isomorphic to the ring of functions on the unipotent subgroup (conjectured in \cite{anderson2003polytope} and proved in \cite{baumann2019mirkovic}). % conjectured by JEA and proved by BKK
% 
To quote from \cite{mirkovic2007geometric}:
\begin{quotation}
    However, we make crucial use of an idea of Drinfeld, going back to around 1990. He discovered an elegant way of obtaining the commutativity constraint by interpreting the convolution product of sheaves as a ``fusion'' product.
\end{quotation}
%
The purpose of this paper is to give a conceptually elementary way to compute this product in type $A$. We do not know how fruitful this work might be in the larger landscape of geometric representation theory as it is plagued by two major limitations: it does not readily generalize outside of type $A$, and it is computationally expensive. 
%

@@

Fix $G = \GL_m$ and let $T\subset G$ be the maximal torus of diagonal matrices. 
% are there other possibilities?? :S 
We identify $\ZZ^m$ with the coweight lattice of $G$ and say that a coweight 
% The coweight lattice of $ G $ is $ \ZZ^m = \Hom(\CC^\times,T)$.
% TODO: Thank you - resolved!
% \acom{This identification assumes we are taking $\CC$ points of $G$ right? i.e.\ it's really $P(T(\CC)) = \Hom(\CC^\times,T(\CC))$ that's $\cong \ZZ^m$ and for instance $P(T(\CC[x]/(x^2))$ will give sth else?} 
% \jcom{No, the coweight lattice is the set of homomorphisms of algebraic groups from $ \mathbb G_m$ to $ GL_m$.  It is independent of the ring/field of definition.}
% 
% The \textbf{dominant coweights} of $ G $ are the set of $ \lambda = (\lambda_1, \dots, \lambda_m) \in \ZZ^m $ such that $ \lambda_1 \ge \cdots \ge \lambda_m$. 
% 
% A coweight $ \mu = (\mu_1, \dots, \mu_m) \in \ZZ^m$ is called \textbf{effective} if $ \mu_j \ge 0 $ for all $ j$.
% 
$\nu = (\nu_1,\dots,\nu_m)$ is \new{dominant} if $\nu_1\ge\cdots\ge\nu_m$ and \new{effective} if $\nu_j\ge 0$ for all $j$.
% We also need the notion of effective dominant coweight, 
% \acom{can we conflate or should we be precise and say coweight?} \jcom{Corrected, let's stick to coweights.}
%which means those $ \lambda $ as above with $ \lambda_m \ge 0$.  
% \acom{1. No boldface on this new term. Proceed with selective boldfacing? 2. Since we sometimes need $\Peff$ which contains $P_{++} = \Peff\cap\Pdom$, how about just using the notation $\Peff,\Pdom$ instead of $P_+, P_{++}$ for a change?}
% \jcom{I added the definition of effective coweight.  I actually don't think that we need any special notation $ P, P_+, P_{dom}$ etc !  Later we can just write ``$ \lambda \in \ZZ^m$ a dominant or effective coweight'' etc}
% TODO: it's settled then, no special notation fro coweights!
% 
If $\nu\in\ZZ^m$ is both effective and dominant then it is a partition of size $|\nu| = \nu_1 + \cdots + \nu_m\in\NN$.
%We write $P_+$ for the set of dominant coweights, and $P_{++}$ for the set of effective dominant coweights. 
% 
% nicematrix test
% $\begin{bNiceArray}{ccc|c}[margin]
%     \Block{3-3}<\Large>{A} & & & 0 \\
%     & \hspace*{1cm} & & \Vdots \\
%     & & & 0 \\
%     \hline
%     0 & \Cdots& 0 & 0
%     \end{bNiceArray}$
% 
% $\begin{pNiceMatrix}
%     \frac12 & -\frac12 \\
%     \frac13 & \frac14 \\
%     \end{pNiceMatrix}$
% 
% Set $\AA = \AA^1_x$ (the subscript denoting the chosen local coordinate) and fix $s\in \AA - \{0\}$. 
% 
% Given a coweight $\mu \in \ZZ^m$ and a point $ s\in \CC$, we define $ (t-s)^\mu$ to be the diagonal matrix 
Given $s\in\CC$ we define $(t-s)^\nu$ to be the diagonal matrix 
% with entries $ (t-s)^{\mu_1}, \dots, (t-s)^{\mu_m}$
% \[
%     \begin{pNiceMatrix}[nullify-dots,xdots/line-style=loosely dotted]
%         (t-s)^{\mu_1} & 0             & \Cdots & 0 \\
%         0             & (t-s)^{\mu_2} & \Ddots & \Vdots \\
%         % 0 & b & a & \Ddots & & \\
%         \Vdots              & \Ddots        & \Ddots & 0 \\
%         % \Vdots & & & & \\
%         0             & \Cdots        &   0    & (t-s)^{\mu_m}
%     \end{pNiceMatrix}
% \]
\[
\begin{bmatrix}
    (t-s)^{\nu_1} \\
    & (t-s)^{\nu_2} \\ 
    & & \ddots \\
    & & & (t-s)^{\nu_m}
\end{bmatrix} 
\]
which we can view in 
% $\GL_m(K) $ 
$G(K)$ for any ring $ K $ containing $(t-s)^{-1}$ and $\CC[t]$. For example, $(t-s)^\nu\in G(K)$ for $ K = \Ks$ or $\CC(t)$. % A: Think R sb K; changed it

\acom{We will also sometimes write $(t-s)^\mu$ for its image in $G(K)/G(R)$ and we hope it will be clear? Only we haven't introduced this space yet.}
% Resolved
% 2. Change notation $\mu = (\mu_1,\dots,\mu_m)$ since we also have $\mu = \mu_1 + \mu_2$. We could do $\mu = \mu' + \mu''$ or $\mu = a + b$? The notation $\mu^{(i)}$ is already taken to mean sth. Ofc all can be changed. Can also do $\mu = {}_{1}\mu + {}_{2}\mu$. \emoji{sweat-smile}}

We will also be interested in the affine space $ M_m$ of $m\times m$ matrices.  Note that for any $ \CC$-algebra $ R $, 
% $\GL_m(R) $ 
$G(R)$ consists of those matrices $ g \in M_m(R) $ whose determinant is invertible in $ R$. Thus for example $ (t-s)^\mu\in M_m(\CC[t])$ for all effective $ \mu \in \ZZ^m$ but 
% $ t^\lambda \in \GL_m(\CC[t]) $ 
$(t-s)^\mu\in G(\CC[t])$ if and only if $ \mu = 0 $.
%

\acom{Maybe add that it's the same for $G(\cO)$. Also, we are using the notation $G(R)$ in one paragraph and defining it in the next. Maybe introduced $M_m$ first? }
